---
title: "Create a Map with ElmerGeo Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create-a-map-with-elmergeo-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE} 
library(psrcelmer)
library(leaflet)
```

With just two packages you can create an interactive map that displays layers from PSRC's *ElmerGeo* database.  This short article will show you how. 

## What is *Leaflet*?

Leaflet is a popular open-source JavaScript library for interactive maps.  R has a package *leaflet* that makes it easy to integrate and control Leaflet maps in R.  Leaflet maps can be made to display the spatial data in one or more *simple-feature* (or *sf*) layers. 


## What is *sf*?

*SF* stands for "simple features", a standardized way of encoding spatial data.  In R, spatial datasets are typically stored as *sf* objects.  

## So how do I make a map?

You can make a simple map in essentially two steps (three if you count loading the libraries).  The example below create a layer of county polygons from the COUNTY_BACKGROUND feature layer in ElmerGeo.


1. Create the layer

Here we create an *sf* layer from the data in ElmerGeo, using the `st_read_elmergeo()` function:
```
  my_county_layer <- st_read_elmergeo('COUNTY_BACKGROUND')
```

2. Display our layer on a leaflet map.  

This creates a map with very simple styling.  You can learn how to enhance its aesthetics through the [leaflet documentation](https://rstudio.github.io/leaflet/).

If you are using RStudio this map should automatically be rendered in the *viewer* pane.
```
  leaflet() %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(data = my_county_layer)
```

